
```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "P:/Shared_documents/Serafina Plüss/20230320_sequencing_results/Janina", cho = TRUE, warning = FALSE, message = FALSE)

# library(ggdendro)
# library(ggrepel)

 #  unloadNamespace("mia")
```

```{r, echo =FALSE}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/metabaRpipe-source/master/Rscripts/functions.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")

#get metadata file: load this function
sample.data.frame <- function(ps) {
  return(as(phyloseq::sample_data(ps), "data.frame"))}

unloadNamespace("mia")
```


## Test Phyloseq
```{r}
phylo_test <-readRDS("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/PB_Inno_Niche/phyloseq_phylo.RDS")
# phylo_test <-readRDS("P:/Shared_documents/Janina_Zuend/08_phd/Raw_Data_PB/phyloseq/phyloseq.RDS")
phylo_test%>%
  physeq_add_metadata(physeq = .,
                      metadata = "P:/Shared_documents/Janina_Zuend/08_phd/Raw_Data_PB/01_metadata_files/metadata_file_TRYPTOPHAN_missing_1.xlsx" %>%
                        readxl::read_xlsx(),
                      sample_column = "sample_name") -> phylo_test


meta_trp<-sample.data.frame(phylo_test)
uplc <- read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/14_working_package_3/01_indoles_quantification/concentrations_calculated_20231121.xlsx" )


left_join(meta_trp, uplc[, c("donor_name", "substrate", "Indolelactate", "Indoleacetate", "Indole", "Indolepropionate", "Tryptamine", "Tryptophan_HN" )])%>%
  gather(metabolite, conc, "Indolelactate": "Tryptamine")%>%
  mutate(conc = ifelse(conc <0.005, 0, conc))%>%
  pivot_wider(names_from = metabolite, values_from = conc)%>%
  left_join(cluster)%>%
  as.data.frame()%>%
  mutate(prod ="no")%>%
  mutate(prod = ifelse(donor_name %in% c("JCDD",  "JBJW",  "GCVJ"), "IPA_only", prod))%>%
  mutate(prod = ifelse(donor_name %in% c("FNNW","BXDM"), "ILA_only", prod))%>%
  mutate(prod = ifelse(donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ"), "IPA_ILA", prod))%>%
  mutate(prod = ifelse(donor_name %in% c("YQRV", "YFKF", "EAJK", "KTRS"), "nothing", prod))->meta_trp_new



colnames(meta_trp_new)[which(names(meta_trp_new) == "Tryptophan_HN")] <- "Tryptophan"

 rownames(meta_trp_new)<- NULL

rownames(meta_trp_new)<-c(meta_trp_new$sample_name)
sample_data(phylo_test)<-meta_trp_new


phylo_test<-phyloseq_get_strains(phylo_test)

noNA = !is.na(tax_table(phylo_test)[,"Genus"]) & !is.na(tax_table(phylo_test)[,"Species"])
tax_table(phylo_test)[noNA][,"Species"] = paste(tax_table(phylo_test)[noNA][,"Genus"], tax_table(phylo_test)[noNA][,"Species"])
phylo_test%>%
  tax_fix()->phylo_test

ps_strain_filt <- phylo_test

# threshold in %
threshold = 0.01

# filter per sample to remove all ASV lower in abundance than threshold (%)
otu_table(ps_strain_filt) <- otu_table(ps_strain_filt) %>%
  as.data.frame() %>%
  dplyr:: mutate(across(everything(), ~ ifelse(. <= sum(.) * (threshold/100), 0, .))) %>% otu_table(., taxa_are_rows = TRUE)

# remove all ASVs that are not present in any sample
ps_strain_filt <- ps_strain_filt %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE)


uplc%>%
  subset(substrate %in% "PE")
```


```{r}
beta <- read.xlsx("C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/beta_distances.xlsx" )
```



##Aldex using Mia
```{r, echo=FALSE, message=FALSE}
#BiocManager::install("microbiome/mia")

 library(mia)
 library(ALDEx2)
```


```{r}
trp <- makeTreeSummarizedExperimentFromPhyloseq(ps_strain_filt%>%subset_samples(!tryptophan %in% "feces")) 
trp_fam <- trp %>% agglomerateByRank(rank = "Family")
trp_gen <- trp %>% agglomerateByRank(rank = "Genus")
trp_species <- trp %>% agglomerateByRank(rank = "Species")


```

###Aldex clusters against each other
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)

  all_trp_gen_sub <-trp_gen[ ,trp_gen$tryptophan %in% c("Trp")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$cluster_hclust)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_trp <- data.frame(x_tt, x_effect)
      
#######
    all_trp_gen_sub <-trp_gen[ ,trp_gen$tryptophan %in% c("no-Trp")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$cluster_hclust)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_no_trp <- data.frame(x_tt, x_effect)
     
```

```{r}
unloadNamespace("mia")

aldex_out_no_trp$tryptophan<-"no-Trp"

aldex_out_no_trp%>%
  rownames_to_column->aldex_out_no_trp

aldex_out_trp$tryptophan<-"Trp"
aldex_out_trp%>%
  rownames_to_column->aldex_out_trp

diff_clusters<-rbind(aldex_out_no_trp, aldex_out_trp )
```

```{r}
diff_clusters%>%
  ggplot(aes(x=diff.btw, y=wi.ep))+
    geom_point(data = subset(diff_clusters, wi.ep < 0.05 ), aes(colour=tryptophan), alpha=0.75, size=3)+
   geom_point(data = subset(diff_clusters, wi.ep > 0.05 ), colour="grey", alpha=0.75, size=3)+
geom_hline(yintercept = 0.05, linetype = "dashed", colour = "red")+
  geom_vline(xintercept = 0.0, linetype = "dashed", colour = "grey")+
  scale_colour_manual(values=c( "darkgoldenrod1", "cyan4"))+
      geom_text_repel(data = subset(diff_clusters, wi.ep <0.05 & (diff.btw > 3 | diff.btw < -3)), aes(label = rowname), size=6, fontface = "italic", max.iter = 10000 )+
 
      theme(axis.text.y= element_text(size = 20), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
            panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
      xlab("Median clr-differences")+ylab("p-value")+scale_y_continuous(
    trans  = compose_trans("log10", "reverse"),
    labels = label_log()
  ) +
      ggtitle("Cluster 1 vs. Cluster 2")

ggsave(filename = "vulcano_genuses.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 17,
       height = 25,
       units = c("cm"))

```

```{r}
diff_clusters%>%filter(wi.ep <0.05)->filter

tax_level ="Genus"

phylo_trp%>%
  physeq_glom_rename(                     speedyseq = T)%>%
   microbiome::transform("clr") %>% 
  psmelt()%>%
    subset(tryptophan %in% c("Trp", "no-Trp"))%>%
   subset(substrate %in% c( "AG", "bGlc", "IN",  "Nut",  "Pea",  "PE", "SS", "XY", "Muc", "H2O"))%>%
  # subset(Genus %in% c(unique(filter$rowname)))%>%
  subset(Genus %in% "Enterocloster")-> catus

# catus %>%
#   # left_join(., cluster)%>%
#   # mutate(Genus = reorder(Genus, Abundance))%>%
#   ggplot(aes(y=Species, x=Abundance))+
#  geom_point(aes(colour=substrate), position = position_dodge(0.9))+
#   geom_boxplot(aes(fill=tryptophan), outlier.shape =NA, alpha=0.5)+
#   # facet_grid(cols=vars(tryptophan))+
#    theme(axis.text.y= element_text(size = 15), 
#               axis.title.y = element_text(size=15), 
#               axis.title.x = element_text(size=15),
#               axis.text.x = element_text(size=15),
#               legend.text = element_text(size = 15),
#               legend.title = element_text(size= 15),
#               plot.title = element_text(size=15), 
#             panel.background = element_rect(fill="white"),
#         panel.border=element_rect(fill=NA), strip.text.y = element_text(angle=0))+
#   ggtitle("Clr-abundances in the two clusters")+
#   facet_grid(rows=vars(donor_name))+
#   scale_fill_manual(values=c( "coral3", "cyan4"))


catus%>%
  subset(!tryptophan %in% "no-Trp")%>%
  ggplot(aes(x= Abundance, y= Indolepropionate))+
  geom_point(aes(colour=substrate))+
  facet_grid(cols=vars(donor_name), rows= vars(OTU))


ggsave(filename = "diffabund_clusterss.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 25,
       height = 25,
       units = c("cm"))
```


##Diff abundance between basal and substrate with Trp -Pea
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)

  all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp") &
                               trp_species$substrate %in% c("Pea", "H2O") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM", "YQRV",  "EAJK")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$substrate)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_substrate <- data.frame(x_tt, x_effect)
      
       aldex_out_substrate$comparison <-"substrate-specififc"
       
      aldex_out_substrate%>%
         rownames_to_column("Species")->aldex_out_substrate
      
      
    ###############################################
all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp", "no-Trp") &
                               trp_species$substrate %in% c("Pea") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM", "YQRV",  "EAJK")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$tryptophan)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_trp <- data.frame(x_tt, x_effect)
       aldex_out_trp$comparison <-"tryptophan-specififc"
       aldex_out_trp%>%
         rownames_to_column("Species")->aldex_out_trp
```

```{r}
rbind(aldex_out_trp[,c("Species","wi.ep","diff.btw", "comparison" )], aldex_out_substrate[,c("Species","wi.ep","diff.btw", "comparison" )])%>%
  mutate(wi.ep = wi.ep *2)%>%
  subset(wi.ep <0.05)->pea
```

##Diff abundance between basal and substrate with Trp -Muc
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)

  all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp") &
                               trp_species$substrate %in% c("Muc", "H2O") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM", "YQRV",  "YFKF", "KTRS")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$substrate)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_substrate <- data.frame(x_tt, x_effect)
      aldex_out_substrate$comparison <-"substrate-specififc"
      aldex_out_substrate%>%
         rownames_to_column("Species")->aldex_out_substrate
    ###############################################
all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp", "no-Trp") &
                               trp_species$substrate %in% c("Muc") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM", "YQRV",  "YFKF", "KTRS")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$tryptophan)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_trp <- data.frame(x_tt, x_effect)
       aldex_out_trp$comparison <-"tryptophan-specififc"
       aldex_out_trp%>%
         rownames_to_column("Species")->aldex_out_trp
```

```{r}
rbind(aldex_out_trp[,c("Species","wi.ep","diff.btw", "comparison" )], aldex_out_substrate[,c("Species","wi.ep","diff.btw", "comparison" )])%>%
   mutate(wi.ep = wi.ep *2)%>%
  subset(wi.ep <0.05)->muc
```

##Diff abundance between basal and substrate with Trp -PE
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)

  all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp") &
                               trp_species$substrate %in% c("PE", "H2O") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$substrate)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_substrate <- data.frame(x_tt, x_effect)
      aldex_out_substrate$comparison <-"substrate-specififc"
      aldex_out_substrate%>%
         rownames_to_column("Species")->aldex_out_substrate
    ###############################################
all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp", "no-Trp") &
                               trp_species$substrate %in% c("Muc") &
                                 trp_species$donor_name %in% c("AVCW", "XNMN", "HTKE", "GWDF", "SEEQ", "JCDD",  "JBJW",  "GCVJ", "FNNW","BXDM")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$tryptophan)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_trp <- data.frame(x_tt, x_effect)
       aldex_out_trp$comparison <-"tryptophan-specififc"
       aldex_out_trp%>%
         rownames_to_column("Species")->aldex_out_trp
```

```{r}
rbind(aldex_out_trp[,c("Species","wi.ep","diff.btw", "comparison" )], aldex_out_substrate[,c("Species","wi.ep","diff.btw", "comparison" )])%>%
   mutate(wi.ep = wi.ep *2)%>%
  subset(wi.ep <0.05)->pe
```
##


##Diff abundance on Pectin
###Aldex clusters against each other
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)

  all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("Trp") &
                               trp_species$substrate %in% c("PE") & 
                           trp_species$prod %in% c("nothing", "IPA_ILA")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$prod)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_trp <- data.frame(x_tt, x_effect)
      
#######
    all_trp_gen_sub <-trp_species[ ,trp_species$tryptophan %in% c("no-Trp")  &
                               trp_species$substrate %in% c("PE") & 
                           trp_species$prod %in% c("nothing", "IPA_ILA")]

       x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$prod)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out_no_trp <- data.frame(x_tt, x_effect)
     
```

```{r}
aldex_out_trp%>%
  subset(wi.ep <0.05)%>%
  mutate(tryptophan = "Trp")->aldex_out_trp

aldex_out_no_trp%>%
  subset(wi.ep < 0.05)%>%
  mutate(tryptophan = "no-Trp")->aldex_out_no_trp

diff_pectin<-rbind(aldex_out_trp, aldex_out_no_trp)%>%
  rownames_to_column("rowname")
```

```{r}
tax %>%
  rownames_to_column("rowname")%>%
  subset(Species %in% "UCG-003 Genus")
   # subset(rowname %in% c("ASV0235", "ASV0165", "ASV0524", "ASV1802", "ASV3197", "ASV3741"))#C.birfementans
  # subset(rowname %in% c("ASV1148", "ASV0867", "ASV0571", "ASV0416"))#C. sporogenes
#subset(rowname %in% c("ASV1148", "ASV0867	"))#C. botulinum
# subset(rowname %in% c("ASV0524"))#p. assacharolyticus
# subset(rowname %in% c("ASV0524", "ASV0165"))
# subset(rowname %in% c("ASV0524", "ASV0165", "ASV0235"))
```


```{r}
 unloadNamespace("mia")


tax <- as.data.frame(tax_table(ps_strain_filt))%>%
  rownames_to_column("rowname")

phylo_trp%>%
  physeq_glom_rename(taxrank = "Strain", 
                     speedyseq = T)%>%
   microbiome::transform("clr") %>% 
  psmelt()%>%
  left_join(., tax)%>%
  # subset(prod %in% c("nothing", "IPA_ILA"))%>%
  subset(tryptophan %in% c("no-Trp","Trp"))%>%
  subset(!substrate %in% c("LT", "FO", "SU"))%>%
   # subset(substrate %in% c( "PE"))%>%
  subset(Genus %in% c("Intestinibacter"))%>%
  mutate(Species = reorder(Species, Abundance))%>%
  ggplot(aes(y=donor_name, x=Abundance))+
 geom_point(aes(colour=substrate, group=tryptophan), position = position_dodge(0.9))+
  geom_boxplot(aes(fill=tryptophan), outlier.shape =NA, alpha=0.5)+
  facet_grid( rows=vars(donor_name), scales = "free")+
  geom_vline(xintercept = 0)+
   theme(axis.text.y= element_text(size = 15), 
              axis.title.y = element_text(size=15), 
              axis.title.x = element_text(size=15),
              axis.text.x = element_text(size=15),
              legend.text = element_text(size = 15),
              legend.title = element_text(size= 15),
              plot.title = element_text(size=15), 
            panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  ggtitle("Clr-abundances in the two clusters")+
  scale_fill_manual(values=c( "coral3", "cyan4", "green3", "darkgoldenrod2", "black"))

ggsave(filename = "diffabund_clusterss.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 30,
       height = 30,
       units = c("cm"))
```

```{r}
ps_strain_filt%>%
  physeq_glom_rename(taxrank = "Species", 
                     speedyseq = T)%>%
   microbiome::transform("clr") %>% 
  psmelt()%>%
  left_join(., tax)%>%
  # subset(prod %in% c("nothing", "IPA_ILA"))%>%
  subset(tryptophan %in% c("Trp", "no-Trp"))%>%
   subset(!substrate %in% c( "FO", "LT", "SU"))%>%
  subset(Species %in% "Intestinibacter bartlettii")%>%
  mutate(Species = reorder(Species, Abundance))%>%
  ggplot(aes(y=donor_name, x=Abundance))+
 geom_point(aes(colour=substrate, group=tryptophan), position = position_dodge(0.9))+
  geom_boxplot(aes(fill=tryptophan), outlier.shape =NA, alpha=0.5)+
   theme(axis.text.y= element_text(size = 15), 
              axis.title.y = element_text(size=15), 
              axis.title.x = element_text(size=15),
              axis.text.x = element_text(size=15),
              legend.text = element_text(size = 15),
              legend.title = element_text(size= 15),
              plot.title = element_text(size=15), 
            panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  ggtitle("Clr-abundances in the two clusters")+
  scale_fill_manual(values=c( "coral3", "cyan4", "green3", "darkgoldenrod2"))+
  facet_grid(cols=vars(tryptophan))

```











###Aldex - for each donor indulvidually => for heatmap
```{r, message=FALSE}
##loop over all donors (are called repsodners anyway ;)
responders<- unique(meta_trp_new$donor_name)
all_trp_gen <-data.frame(matrix(ncol=14, nrow=0))

for (i in (1:length(responders))){
  all_trp_gen_sub <-trp_gen[ ,trp_gen$donor_name %in% c(responders[i])]

 
    result_name <- paste(responders[i])
   
      x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$tryptophan)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name

      names(result)[6]<-paste0("rab.win.no")
      names(result)[7]<-paste0("rab.win.Trp")
      result$genus<-row.names(result)
      all_trp_gen<-rbind(all_trp_gen, result)
}

all_trp_gen ->out

all_trp_gen%>%
  filter(we.ep < 0.05)%>%
  filter(rab.win.no > 5| rab.win.Trp >5)->sig
```
## Vulcano plot
```{r}
# library(ggrepel)
out%>%
ggplot(aes(x=diff.btw, y=we.ep))+
  geom_point(data = subset(out, we.ep < 0.05 ), colour="cyan4", alpha=0.75, size=3)+
  geom_point(data = subset(out, we.ep > 0.05 ), colour="grey", alpha=0.75, size=3)+
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "red")+
  geom_vline(xintercept = 0.0, linetype = "dashed", colour = "grey")+
      geom_text_repel(data = subset(out, we.ep < 0.05 ), aes(label = genus), fontface="italic", size=5)+
      theme(axis.text.y= element_text(size = 20), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
            panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
      xlab("Median clr-differences")+ylab("p-value")+scale_y_continuous(
    trans  = compose_trans("log10", "reverse"),
    labels = label_log()
  ) +
      ggtitle("Trp-treated vs. controls (responders)")+
  facet_grid(cols=vars(V14))

```



## Heatmap & with donor clustering using hclust
```{r}
unloadNamespace("mia")
out[, c("V14", "genus", "diff.btw")]%>%
  subset(genus %in% unique(sig$genus))->mat

rownames(mat)<-NULL

####hclust donornames
mat%>%
  pivot_wider(values_from = diff.btw, names_from = genus) ->sub

sub%>%
  column_to_rownames(var = 'V14') %>% 
  as.matrix()->df_diff
diff_dendro <- as.dendrogram(hclust(d = dist(x = df_diff)))
# Create dendro
dendro_plot_donor <- ggdendrogram(data = diff_dendro, rotate = F)
dendro_plot_donor
diff_order <- order.dendrogram(diff_dendro)

out%>%
  subset(genus %in% unique(sig$genus))->heat_donor
  
heat_donor%>%
  left_join(., setNames(meta_trp_new[, c("donor_name",  "cluster_hclust")], c("V14", "cluster_hclust")))%>%
   mutate(genus = reorder(genus, diff.btw))->sub

sub%>%
  ggplot(aes(y=genus, x=V14, fill=diff.btw))+
  geom_tile()+
  geom_text(data = subset(sub, we.ep < 0.05 ), label="*")+
   scale_fill_gradientn(colors = c( "coral", "white", "cyan4"),  na.value= "grey", values = rescale(c(min(heat_donor$diff.btw) , 0, max(heat_donor$diff.btw))))+
  theme(axis.text.y= element_text(size=15), 
              axis.title.y = element_blank(), 
              axis.title.x = element_blank(),
              axis.text.x = element_text(size=15, angle=90),
              legend.text = element_text(size = 15),
              legend.title = element_text(size= 15),
              plot.title = element_text(size=15), 
            panel.background = element_rect(fill="gray93"),
        panel.border=element_rect(fill=NA),
        panel.grid.major = element_blank(),
  panel.grid.minor = element_blank())+
  facet_grid(cols=vars(cluster_hclust), scales="free", space="free")+
  ggtitle("Differential abundance - Trp vs. control")

ggsave(filename = "heat.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 20,
       height = 19,
       units = c("cm"))

```

### aldex based on substarte
```{r}
# responders <- c("YFKF",  "EAJK", "KTRS", "TPSJ", "HXBH" ,  "JCDD", "GCVJ", "YQRV")
responders<- c( "AG"   , "bGlc" , "IN"   , "Muc"  , "Nut"   ,"Pea"   ,"PE"   , "SS"  , "XY"  ,  "H2O" )
result_sub <-data.frame(matrix(ncol=14, nrow=0))

for (i in (1:length(responders))){
  all_trp_gen_sub <-trp_gen[ ,trp_gen$substrate %in% c(responders[i])]

 
    result_name <- paste(responders[i])
   
      x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$tryptophan)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name

      names(result)[6]<-paste0("rab.win.no")
      names(result)[7]<-paste0("rab.win.Trp")
      result$genus<-row.names(result)
      result_sub<-rbind(result_sub, result)
  }


result_sub ->out_substrate

result_sub%>%
  filter(we.ep < 0.05)->sig_substrate

```

```{r}
out_substrate[, c("V14", "genus", "diff.btw")]%>%
  subset(genus %in% unique(sig_substrate$genus))->mat

rownames(mat)<-NULL

####hclust donornames
mat%>%
  pivot_wider(values_from = diff.btw, names_from = genus) ->sub

sub%>%
  column_to_rownames(var = 'V14') %>% 
  as.matrix()->df_diff
diff_dendro <- as.dendrogram(hclust(d = dist(x = df_diff)))
# Create dendro
dendro_plot_donor <- ggdendrogram(data = diff_dendro, rotate = F)
dendro_plot_donor
diff_order <- order.dendrogram(diff_dendro)



out_substrate%>%
  subset(genus %in% unique(sig_substrate$genus))->heat_sub

heat_sub%>%
   mutate(V14 =  factor(V14, levels = sub$`V14`[diff_order], ordered = TRUE),
          genus = reorder(genus, diff.btw))%>%
  ggplot(aes(y=genus, x=V14, fill=diff.btw))+
  geom_tile()+
  geom_text(data = subset(heat_sub, we.ep < 0.05 ), label="*")+
   scale_fill_gradientn(colors = c( "coral", "white", "cyan4"),  na.value= "grey", values = rescale(c(min(heat_sub$diff.btw) , 0, max(heat_sub$diff.btw))))+
  theme(axis.text.y= element_text(size=15), 
              axis.title.y = element_blank(), 
              axis.title.x = element_blank(),
              axis.text.x = element_text(size=15, angle=90),
              legend.text = element_text(size = 15),
              legend.title = element_text(size= 15),
              plot.title = element_text(size=15), 
            panel.background = element_rect(fill="gray93"),
        panel.border=element_rect(fill=NA),
        panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
        )

ggsave(filename = "heat_substrate.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 20,
       height = 14,
       units = c("cm"))
  
```
###Analyse non-responders vs. responders

```{r}
all_trp_gen_sub <-trp_gen[ ,trp_gen$responder_group %in% c("N1", "R1") & trp_gen$tryptophan %in% c("no-Trp")]

   
      x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$responder_group)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
     
aldex_out%>%
  filter(we.ep < 0.05)->sig_1



```
```{r}
all_trp_gen_sub <-trp_gen[ , trp_gen$tryptophan %in% c("no-Trp")]

   
      x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$responder)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
     
aldex_out%>%
  filter(we.ep < 0.05)->sig_2



```

```{r}
all_trp_gen_sub <-trp_gen[ , trp_gen$tryptophan %in% c("no-Trp") &  trp_gen$responder_group %in% c("R1", "R2")]

   
      x <- aldex.clr(
      reads = assay(all_trp_gen_sub),
     all_trp_gen_sub$responder_group)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
     
aldex_out%>%
  filter(we.ep < 0.05)->sig_3
```

```{r}
unloadNamespace("mia")

```

