```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "P:/Shared_documents/Serafina Plüss/20230320_sequencing_results/Janina", cho = TRUE, warning = FALSE, message = FALSE)
# devtools::install_github("tidyverse/tidyverse")
library(tidyverse)
library(openxlsx)
library(ggh4x)
library(microViz)
library(ggpmisc)
library(ggpubr)
library(rstatix)
library(scales)
```

```{r, echo =FALSE}
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_taxa_tests.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_normalisation.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_alpha.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_beta.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_heatmap.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_taxa_tests.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_varia.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_beta.R")
source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/functions.R")
source("https://raw.githubusercontent.com/fconstancias/metabarcodingRpipeline/dev/scripts/functions_export_simplified.R")

#get metadata file: load this function
sample.data.frame <- function(ps) {
  return(as(phyloseq::sample_data(ps), "data.frame"))}

facet_labeller <- function(variable,value){
  return(names[value])}
```

## Load Phyloseq
#this phyloseq already contains metabolic data in the metadata table @ we have to first link the metabolites to the metadata
```{r}
Trp_rare <-readRDS("C:/Users/zuendj/Desktop/03_data R/wp3/input/Trp_rare.RDS")
```

```{r}
meta_trp<-sample.data.frame(Trp_rare)
```

##Corelation of abundances and Trp production
```{r}  
## get the clr abundance into a new data frame
##taxrank to be defined
Trp_rare%>%
  physeq_glom_rename(
    # taxrank="Species",
                     speedyseq = T)%>% 
  microbiome::transform("clr") %>%
  subset_samples(tryptophan %in% c("Trp"))%>%
  psmelt()->df_abund 


#get the mean abundace per donor and OTU for filtering @bp - this you will not do likely?! 
setNames(aggregate(df_abund$Abundance, by=list(df_abund$OTU, df_abund$donor_name), FUN=mean), c("OTU", "donor_name", "mean"))%>%
  # subset(OTU %in% "[Eubacterium] eligens group")
  subset(mean>0)->filter

```
## Corelation analysis
```{r}
#check each donor individually => donor-individual correlation

##step 1. create a data frame for results
correlations<-as.data.frame(matrix(0,0,5))
colnames(correlations)<-c("OTU", "donor_name", "p_value", "corr", "metabolite")


#step 2. define all donors & metabolites to be looped over 
donor_cond<- c("SEEQ", "XNMN", "JCDD", "GWDF", "HTKE", "JBJW", "AVCW", "BXDM", "KTRS", "EAJK", "YQRV", "TPSJ", "HXBH", "GCVJ", "YFKF", "FNNW")
indole_cond<-c("Indole", "Indolepropionate", "Indolelactate", "Tryptamine", "Indoleacetate")
n=1

# step 3. loop over all donors & metabolites & ASVs
for (i in 1:length(donor_cond)){
  
#subset donor => @bp - this you will not do likely?!
filter%>%
    subset(donor_name %in% donor_cond[i])->filt_don
#filter  
df_abund %>%
  subset(OTU %in% filt_don$OTU)->df_filt

#subset donor => @bp - this you will not do likely?!
  df_filt%>%
    subset(donor_name %in% donor_cond[i])->df_donor
  

otu_cond<-unique(df_donor$OTU)

for (j in 1:length(otu_cond)){ #loop over all OTUs
  for(x in 1:length(indole_cond)){
    
    df_donor%>% 
      subset(OTU %in% c(otu_cond[j]))-> corr
    
    ### corraltion calculation - put results into a new variable
   result<-cor.test( corr[,indole_cond[x]] , corr$Abundance)
   
   
#store all results and conditions etc. in a new vector 
out<-c(otu_cond[j], donor_cond[i], result$p.value, as.data.frame(result$estimate)[1,1], indole_cond[x])

#add result vector to result data frame
correlations[n,]<-out

n=n+1

  }}}

```
##Filtering
## i added some filtering steps: only significant correlations with metabolite concentrations that were higher tha 0.03 mM were further analyzed @bp you can skip
```{r}

df_abund %>%
   gather(metabolite, conc, "Tryptophan":"Tryptamine")->df_long

setNames(aggregate(df_long$conc, by=list(df_long$donor_name, df_long$metabolite), FUN="max"), c("donor_name", "metabolite", "conc"))

left_join(correlations, setNames(aggregate(df_long$conc, by=list(df_long$donor_name, df_long$metabolite), FUN="max"), c("donor_name", "metabolite", "conc")))%>%
  mutate(p_value = ifelse((conc > 0 & conc <0.03), 0.06, p_value))-> correlations_conc_filtered ### that's the filtered data frame

correlations_conc_filtered->corr_donor
```

## Figure
```{r}
# thats a function to label the facets of the ggplot nicely @bp you can skip
facet_labeller <- function(variable,value){
  return(names[value])}


#names for the facets of the ggplot @bp you can skip
names <- list("Indolepropionate"= "IPA",
              "Indolelactate"="ILA", 
              "Indoleacetate"="IA",
              "Indole"= "Indole", 
              "Tryptamine"="Tryptamine")

#selection of significant correlations & only correlations that were found in more that 3 donors => @bp you can skip
correlations_conc_filtered%>%
  mutate(corr = as.numeric(corr),
         p_value=as.numeric(p_value))%>%
  subset(p_value <0.05 & corr > 0.5)%>%
  group_by( metabolite, OTU) %>% # Group by donor_name, metabolite, and OTU
  filter(n() >= 3) %>% # Keep only groups with at least 3 entries
  ungroup()->fil_corr




# link the tax table so that we have full taxonomic classification
left_join(correlations_conc_filtered, as.data.frame(tax_table(Trp_rare)) %>% 
            rownames_to_column("OTU"))%>%
   subset(OTU %in% c(fil_corr$OTU))%>%
  mutate(corr = as.numeric(corr),
         p_value=as.numeric(p_value))%>%
   subset(p_value < 0.05)->sig

frequency <- sig$Species %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Species", "freq"))


left_join(sig, frequency)%>%
  mutate(metabolite = factor(metabolite, levels=c("Indolepropionate", "Indolelactate", "Indoleacetate", "Indole", "Tryptamine")))%>%
  mutate(Species = ifelse(Species == "Lachnospiraceae NK4A136 group bacterium", "Lachnospiraceae NK4A136", Species))%>% # @bp you can skip
  mutate(Species = ifelse(Species == "Faecalibacterium unknown", "Faecalibacterium", Species))%>% #@bp you can skip
   mutate(Species = reorder(Species, freq))%>% # order base on the number of frequency of correlation fo one specific OTU
 
#change names in case you want @bp you can skip
  mutate(donor_name = ifelse(donor_name == "XNMN", "D1", donor_name))%>%
   mutate(donor_name = ifelse(donor_name == "SEEQ", "D2", donor_name))%>%
   mutate(donor_name =                 ifelse(donor_name == "HTKE", "D3",donor_name))%>%
   mutate(donor_name =                 ifelse(donor_name == "GWDF", "D4",donor_name))%>%
    mutate(donor_name =                ifelse(donor_name == "JCDD", "D5",donor_name))%>%
    mutate(donor_name =                ifelse(donor_name == "BXDM", "D6",donor_name))%>%
     mutate(donor_name =               ifelse(donor_name == "JBJW", "D7",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "AVCW", "D8",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "FNNW", "D9",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "GCVJ", "D10",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "HXBH", "D11",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "TPSJ", "D12",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "YFKF", "D13",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "EAJK", "D14",donor_name))%>%
        mutate(donor_name =            ifelse(donor_name == "KTRS", "D15",donor_name))%>%
        mutate(donor_name =            ifelse(donor_name == "YQRV", "D16", donor_name))%>%

    mutate(donor_name = factor(donor_name, levels =c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14", "D15", "D16")
))%>%
  
  ### let's do a plot !! 
  ggplot(aes(x=donor_name, y=Species))+
  geom_point(aes(fill=corr, size=p_value),colour="black", alpha=0.75,  shape=21, width=0.1, height=0.1)+ 
  facet_grid(col=vars(metabolite),  space="free", scales="free", labeller=facet_labeller)+
  theme(text = element_text(size=15),
          axis.text.x = element_text(angle = 90, vjust=0.5),
        axis.text.y= element_text(face="italic"),
        axis.ticks.y = element_blank(),
        axis.title=element_blank(),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA),
        # legend.position = "bottom",
        legend.title = element_text(size=15),
        # strip.text.y= element_blank(),
        # panel.spacing = unit(0, "lines"),
        strip.background = element_blank())+
  scale_size_continuous(range = c(5, 1), name="P-value") +
  scale_fill_gradientn(colors = c("indianred",  "white", "cyan4"), values = rescale(c( -1, 0, 1)), name="R")+
  geom_hline(yintercept = c(1.5,2.5,3.5, 4.5, 6.5, 7.5,8.5, 9.5, 10.5, 11.5, 12.5, 13.5, 14.5, 15.5, 16.5),  colour="grey", linetype="dashed", alpha=.5) # @bp you can skip


ggsave(filename = "correlating_taxa_responders_ASVs..jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 25,
       height = 10,
       units = c("cm"))
```


# highlight correaltion for specififc taxa & donors?? @bp you can skip

```{r}
Trp_rare%>%
  physeq_glom_rename( speedyseq = T)%>%
  microbiome::transform("clr") %>% 
  psmelt()%>%
    mutate(donor_name = ifelse(donor_name == "XNMN", "D1", donor_name))%>%
   mutate(donor_name = ifelse(donor_name == "SEEQ", "D2", donor_name))%>%
   mutate(donor_name =                 ifelse(donor_name == "HTKE", "D3",donor_name))%>%
   mutate(donor_name =                 ifelse(donor_name == "GWDF", "D4",donor_name))%>%
    mutate(donor_name =                ifelse(donor_name == "JCDD", "D5",donor_name))%>%
    mutate(donor_name =                ifelse(donor_name == "BXDM", "D6",donor_name))%>%
     mutate(donor_name =               ifelse(donor_name == "JBJW", "D7",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "AVCW", "D8",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "FNNW", "D9",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "GCVJ", "D10",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "HXBH", "D11",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "TPSJ", "D12",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "YFKF", "D13",donor_name))%>%
       mutate(donor_name =             ifelse(donor_name == "EAJK", "D14",donor_name))%>%
        mutate(donor_name =            ifelse(donor_name == "KTRS", "D15",donor_name))%>%
        mutate(donor_name =            ifelse(donor_name == "YQRV", "D16", donor_name))%>%

    mutate(donor_name = factor(donor_name, levels =c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14", "D15", "D16")
))%>%
  subset(OTU %in% sig$OTU)%>%
  subset(tryptophan %in% "Trp")%>%
  subset(Species %in% c( "Lachnospira pectinoschiza"))%>%
  # subset(Genus %in% c("[Eubacterium] eligens group"))%>%
  mutate(reductive_pathway = 1000 * (Indolepropionate + Indolelactate))%>%
  mutate(substrate = factor(substrate, levels=c( "H2O","SS"  , "XY" ,  "Nut" , "AG"  ,   "IN"  , "bGlc" ,"PE" ,  "Pea")))%>%
  ggplot(aes(y=Abundance, x=reductive_pathway))+
  # geom_jitter(aes(colour=donor_name))+
  # geom_boxplot()+
  geom_hline(yintercept = 0, linetype="dashed", colour="grey")+
  geom_vline(xintercept = 0.0, linetype="dashed", colour="grey")+
  
  
  geom_point(aes(colour=substrate, shape =OTU))+
  stat_poly_line(color="darkgrey") +
  stat_correlation(mapping = use_label(c("R", "P")), size=4)+
  facet_wrap(vars(donor_name))+
  xlab(expression(Reductive~metabolites~"["*mu*"M]"))+ylab("Clr-abundance")+labs(shape="ASV")+
   scale_colour_manual(values  =  c("grey","#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f"), name= "Substrate", labels = c("H2O"=expression(paste(H[2]*O))))+
  theme(text = element_text(size=15) ,
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA),
        strip.background = element_blank(),
        legend.text = element_text(hjust=0),
        axis.text.x =  element_text(angle=90),
        plot.title = element_text(face="italic"))+
   ggtitle("Lachnospira pectinoschiza")
  # ggtitle("[Eubacterium] eligens group")

ggsave(filename = "pectino_taxa.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp3/InnoNiche_data/output",
       width = 21,
       height = 15,
       units = c("cm"))

```

## Differential Abundance Pectin

```{r, echo=FALSE}
library(mia)
library(miaViz)
library(ALDEx2)
```



```{r}

Trp_filtered <-readRDS("C:/Users/zuendj/Desktop/03_data R/wp3/input/Trp_filtered.RDS")
tse_C <- enetros %>% agglomerateByRank(rank = "Family")
```

###Aldex: NO Prevalence filter-> filter later with abundance
```{r}

substrates <-c("7C",
               "3C+Muc",
               "7C-Muc" ,
               "3C" ,
               "SS" ,
               "Nut" ,
               "Glc" ,
               "H2O" )


all_C <-data.frame(matrix(ncol=14, nrow=0))

for (i in (1:length(substrates))){
  tse_substrates <-tse_C[ , tse_C$condition %in% c(substrates[i], "feces")]
  
      result_name <- paste(substrates[i])
      print(result_name)
      
      count <- as.vector(tse_substrates$condition) 
      x <- aldex.clr(
      reads = assay(tse_substrates),
     tse_substrates$condition)
    # calculates expected values of the Welch's t-test and Wilcoxon rank test on
    # the data returned by aldex.clr
      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      # determines the median clr abundance of the feature in all samples and in
      # groups, the median difference between the two groups, the median variation
      # within each group and the effect size, which is the median of the ratio
      # of the between group difference and the larger of the variance within groups
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      # combine all outputs 
      
      aldex_out <- data.frame(x_tt, x_effect)
      
      
      
      result <-aldex_out
      result[,14] <- result_name
      print(colnames(result))
      
      names(result)[6]<-paste0("rab.win.feces")
      names(result)[7]<-paste0("rab.win.cultures")
      result$Family<-row.names(result)
      
      all_C <- rbind(all_C, result)
                                    
      }

all_C %>%
  subset(Family %in% "Family:Enterobacteriaceae")%>%
  filter(we.ep < 0.05)->out
```

```{r}
unloadNamespace("miaViz")
unloadNamespace("mia")
```

































